Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-11 results.tex
explanation of col arima cluster spread

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. $MASE$ plots for all 360
experiments.
\begin{figure}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=\columnwidth]{figs/predictions_vs_entropy}
\caption{WPE vs. $MASE$ for all trials, methods, and systems. \svdone,
  \svdthree, and \svdfive are omitted from the top plot for scale
  reasons. {\color{red} All three plots need to have the same
    horizontal scale!}
% 
% ; their $MASE$ scores are $2.676 \pm 4.328$, $31.39 \pm 0.282$, and
%   $20.87 \pm 0.192$ respectively. 
% 
Numerical values, including means and standard deviations of the
errors, can be found in Table~\ref{tab:error}.  The dashed line is the
same as in the previous figure.
% 
%  [[All traces except those from \svd$_1$ lie above the line,
%      indicating that LMA is better suited prediction method for the
%      computer traces considered.]]  [[This comes {\bf long} before the
%      associated discussion in the text, and it's not what this figure
%      is trying to show here.]]
% 
}
    \label{fig:wpe_vs_mase_all}
\end{figure} 
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding system in any of the three plots in
Figure~\ref{fig:wpe_vs_mase_all}.)  The WPE values do not vary very
much across trials.  For most traces, the variance in $MASE$ scores is
low as well, resulting in small, tight clusters.  In some
cases---ARIMA predictions of {\tt col\_major}, for instance---the
$MASE$ variance is larger, which spreads out the clusters
horizontally.  The distribution of the cluster {\sl positions} is
quite different in the three plots.  The mean $MASE$ scores of
predictions generated with the nonlinear LMA method are generally
closer to the dashed line; the ARIMA method clusters are more
widely spread, the \naive ~clusters even more so.  This geometric
validation of the second conjecture in this paper is discussed in more
details in the following paragraphs.

@@

The dashed line in the Figures is a heuristic that captures the
argument at the end of the previous paragraph.  It can allow
practitioners to recognize when a prediction method is not well
matched to the task at hand: that is, when the time series has more
predictive structure than the method is able to use.  When the
predictions are well below that line (as in the top and middle images
in Figure~\ref{fig:wpe_vs_mase_all}), one should try a different
method.

Note that the \col points (blue {\color{blue}$+$} icons) are clustered
very tightly in the lower left quadrant of the LMA and \naive ~$MASE$
vs. WPE plots, but spread out horizontally in the ARIMA plot.
{\color{red}[[We need to come up with a hypothesis about this spread,
      even if it's totally tentative.  Was it that the random-walk got
      really good in a few of the \col trials?  Or that ARIMA got
      really bad in those trials?  In either case, why did that
      happen---what was it about those trials that made ARIMA bark
      and/or random walk do well?]]} The position of the LMA
cluster---well to the left of the other two, near the dashed
line---reflects this method's ability to capture and exploit the
structure that is present in this signal.

>>>>

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. $MASE$ plots for all 360
experiments.
\begin{figure}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=\columnwidth]{figs/predictions_vs_entropy3}
\caption{WPE vs. $MASE$ for all trials, methods, and systems. \svdone,
  \svdthree, and \svdfive are omitted from the top plot for scale
  reasons.
% 
% ; their $MASE$ scores are $2.676 \pm 4.328$, $31.39 \pm 0.282$, and
%   $20.87 \pm 0.192$ respectively. 
% 
Numerical values, including means and standard deviations of the
errors, can be found in Table~\ref{tab:error}.  The dashed line is the
same as in the previous figure.
% 
%  [[All traces except those from \svd$_1$ lie above the line,
%      indicating that LMA is better suited prediction method for the
%      computer traces considered.]]  [[This comes {\bf long} before the
%      associated discussion in the text, and it's not what this figure
%      is trying to show here.]]
% 
}
    \label{fig:wpe_vs_mase_all}
\end{figure} 
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding system in any of the three plots in
Figure~\ref{fig:wpe_vs_mase_all}.)  The WPE values do not vary very
much across trials.  For most traces, the variance in $MASE$ scores is
low as well, resulting in small, tight clusters.  In some
cases---ARIMA predictions of {\tt col\_major}, for instance---the
$MASE$ variance is larger, which spreads out the clusters
horizontally.  The distribution of the cluster {\sl positions} is
quite different in the three plots.  The mean $MASE$ scores of
predictions generated with the nonlinear LMA method are generally
closer to the dashed line; the ARIMA method clusters are more
widely spread, the \naive ~clusters even more so.  This geometric
validation of the second conjecture in this paper is discussed in more
details in the following paragraphs.

@@

The dashed line in the Figures is a heuristic that captures the
argument at the end of the previous paragraph.  It can allow
practitioners to recognize when a prediction method is not well
matched to the task at hand: that is, when the time series has more
predictive structure than the method is able to use.  When the
predictions are well below that line (as in the top and middle images
in Figure~\ref{fig:wpe_vs_mase_all}), one should try a different
method.  The position of the LMA cluster, for instance---well to the
left of the other two, near the dashed line---reflects this method's
ability to capture and exploit the structure that is present in this
signal.

Note that the \col points (blue {\color{blue}$+$} icons) are clustered
very tightly in the lower left quadrant of the LMA and \naive ~$MASE$
vs. WPE plots, but spread out horizontally in the ARIMA plot.  This is
because of the way that these models are built.  If a KPSS test of the
time series indicates that it is nonstationary, an integration term is
added to the model.  According to this test, the \col time series is
right on the edge of nonstationarity, so the ARIMA procedure sometimes
omits the integration term, which increases the error, thereby
spreading out the points.
